First steps towards a formalization of Forcing

### 2018-11-17 LSFA/related.tex
formalización de matemática en general

\subsection{Related work}\label{sec:related-work}
Before we proceed, we discuss very succinctly recent formalizations of
set theory and forcing. The closest formalizations are those based on
Isabelle. Let us remark that Isabelle allows for different logical
foundations; in particular, Paulson carried out his formalizations
on top of Isabelle/FOL which is based on first order logic.

>>>>

\subsection{Related work}\label{sec:related-work}
Formalization of mathematics serves many purposes. The most obvious
one is to increase reliability in a result and/or its proof. This has
been the original motivation that lead Voevodsky to gather many
researchers around \emph{homotopy type theory} and its formalization in Coq
\cite{hottbook}; the same applies to the four colour theorem (checked by
Gonthier \cite{MR2463991}) and the formidable \emph{Flyspeck} project
\cite{MR3659768} by the team conducted by Hales.

In our particular case, forcing and the set theoretic techniques that
are being formalized can be regarded as a mature technology and thus
the main goal is not to increase confidence. Nevertheless, the level
of detail in a formalization of this sort always provide additional
information about the inner workings of the theory: It is expected, for
instance, to have a detailed account of which axioms are necessary to
define and use forcing. Finally, we support the vision that a growing
corpus of formalized mathematics can be a useful library for the
future generations. The question of how to systematize this corpus is
an ongoing project by Paulson \cite{ALEXANDRIA}.

We will now discuss very succinctly recent formalizations of
set theory and forcing. The closest formalizations are those based on
Isabelle. Let us remark that Isabelle allows for different logical
foundations; in particular, Paulson carried out his formalizations
on top of Isabelle/FOL which is based on first order logic.
